\documentclass[12pt]{article}
\usepackage{fullpage,amsmath,amsfonts}

\begin{document}

\title{Stochastic Methods + Lab}
\author{Session 12}
\date{October 11, 2016}
\maketitle

\begin{enumerate}


\item Plot the binomial distribution 
\[
  b(j,n;p) = p^j \, q^{n-j} \, \binom{n}{j}
\]
where $q=1-p$ into a coordinate system where the values on the
$x$-axis correspond to $j$ according to
\[
  x_j = \frac{j-np}{\sqrt{npq}}
\]
and the $y$-values are given by $\sqrt{npq}\, b(j,n;p)$.

Compare the graphs for $n=10$, $n=100$, and the graph of the standard
Gaussian
\[
  \phi(x) = \frac1{\sqrt{2\pi}} \, \mathrm{e}^{-x^2/2}
\]
in the same plot.  Comment \emph{briefly} on what you see.


\item Generate $N=10\,000$ samples of the binomial distribution
(number of successes in $n$ independent trials).  Rescale the samples
via
\[
  X = \frac{J- \mathbb E[J]}{\sqrt{\operatorname{Var}[J]}}
\]
where you use the sample mean to approximate $\mathbb E[J]$ and the
sample variance to approximate $\sqrt{\operatorname{Var}[J]}$.  (These
can be computed via the \textsf{numpy}-functions \verb+mean()+ and
$\verb+std()+.$)

Generate $N=10\,000$ samples of the standard normal distribution.

Plot the sorted samples for $X$ vs.\ the sorted samples for the
standard normal distribution.  Comment briefly on what you see.  

\emph{Note:} This is called a QQ-plot and is more generally used to
empirically compare two probability distributions.

\item Compute an ensemble of standard Brownian paths $W(t)$ over the
interval $[0,1]$ partitioned into $N=500$ time steps.  Plot the
empirically determied mean and standard deviation of the ensemble as a
function of time.

\item Similarly, compute an ensemble of geometric Brownian paths
\[
  S(t) = \exp((\mu - \tfrac12 \, \sigma^2)\, t + \sigma \, W(t))
\]
with $\mu=0.05$ and $\sigma=0.3$ and plot mean and standard deviation as
a function of time on the interval $[0,1]$.

\end{enumerate}



\end{document}

