\documentclass[12pt]{article}
\usepackage{fullpage,amsmath,amsfonts}

\begin{document}

\title{Stochastic Methods + Lab}
\author{Session 6}
\date{September 20, 2016}
\maketitle

\noindent\emph{Notes:} 
\begin{itemize}
\item The computational work is to be submitted up to
one week after the session via \texttt{git}.  
\item For this task sheet, it
is sufficient to submit the code so long as it runs and produces the
requested output. 

\item Please do not submit \textsf{ipython} notebooks.  If you prefer
working in the ipython shell, you can convert the notebook to a
standalone script writing
\begin{verbatim}
  ipython3 nbconvert --to python filename.ipynb
\end{verbatim}
or 
\begin{verbatim}
  jupyter nbconvert --to python filename.ipynb
\end{verbatim}
(depending on your installation).  Make sure the file runs; you may
have to add print statements expressions resulting from output cells.

\item Make sure that your submission runs the requested test case.  

\item The theoretical questions can be submitted handwritten on paper.
\end{itemize}

\begin{enumerate}
\item Consider a 10-year level coupon bond with an annual coupon rate
of 8\% compounded annually and a par value of EUR 1\,000.  Plot the
price of the bond vs.\ the yield.

\item Plot price volatility vs.\ time to maturity for level coupon
bonds with annual coupon rates of 2\%, 6\%, and 12\% paid semiannually.
Assume a yield of 6\% and a par value of EUR 1\,000.  To see the
different volatility behaviors, take a range from 0 up to 100 years to
maturity. 

\item Plot the bond value (forward value) of an 8\% 15-year bond
compounded semi-annually vs.\ years to maturity under three rate
scenarios: (a) the interest rate decreases instantaneously to 6\%, (b)
the interest rate remains unchanged, and (c) the interest rate
increases instantaneously to 10\%.

\emph{Note:} The forward value of the bond required here is the value
of the bond at time zero based on the new market interest rate times
the compounded interest factors corresponding to different times to
maturity.

\item Plot the future value of a 30-year bond at a coupon rate of 10\%
compounded annually after a 10-year horizon as a function of yield.
Find the minimum of the horizon price numerically, e.g., using
\texttt{scipy.optimize.brent}. 

\item Visit the web site of the European Central Bank (ECB) and look
for their yield curve data.  What data is contained in the files?

(This is to be submitted as a written answer!)

\item The ECB publishes spot rate data in a downloadable CSV file.
Find online documentation on how to read CSV files, read in the ECB
spot rates, and plot the yield curve.

\item Write out a consistent set of expressions for the Macaulay
duration and the immunization of a level coupon bond assuming
continuous compounding of interest.

\item Suppose you want to immunize a liability at time $D$ with
two bonds of Macaulay durations $D_1$ and $D_2$, respectively.  Show
that the fraction $\omega_1$ and $\omega_2$ of the two bonds in the
initial portfolio need to satisfy
\begin{gather*}
  \omega_1 + \omega_2 = 1 \,, \\
  \omega_1 \, D_1 + \omega_2 \, D_2 = D \,.
\end{gather*}

\end{enumerate}


\end{document}

